Self Referencing Sequences

### 2014-11-01 draft.tex
edited up to prefixes, added a lemma showing sequences are unique

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number. To see why this is true, we give an algorithm to build the sequence. Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$. If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

@@

Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number. From this point on, we will denote by $s(G,g)$ the unique sequence generated by set $G$ with starting number $g \in G$.

>>>>

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.

\begin{lemma}
Let $C$ be a generating cycle. Suppose $(a_i)$ and $(b_i)$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
\end{lemma}

\begin{proof}
Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $(a_i)$ and $(b_i)$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $(a_i)$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $(b_i)$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.

\begin{figure}{hb}
\label{fig:unequalsequences}
\center
\includegraphics[width=0.5\linewidth]{lemma1.pdf}
\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
\end{figure}

This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
\end{proof}

We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1,a_2,\dots,a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

@@

From this point on, we will denote by $s(C,a)$ the unique sequence with generating cycle $C$ and with starting number $a \in C$.
